Weighted Model Counting Without Parameter Variables

### 2021-03-14 doc/paper3/paper.tex
Some editing

% The abstract should briefly summarize the contents of the paper in
% 150--250 words.
% 1. Introduction. In one sentence, what’s the topic?
% 2. State the problem you tackle.
% 3. Summarize (in one sentence) why nobody else has adequately answered the research question yet.
% 4. Explain, in one sentence, how you tackled the research question.
% 5. In one sentence, how did you go about doing the research that follows from your big idea.
% 6. As a single sentence, what’s the key impact of your research?
\begin{abstract}
  Weighted model counting (WMC) is a powerful computational technique for a
  variety of problems, especially commonly used for probabilistic inference.
  However, the standard definition of WMC that puts weights on literals often
  necessitates WMC encodings to include additional variables and clauses just so
  each weight can be attached to a literal. This paper complements previous work
  by considering WMC instances in their full generality and using recent
  state-of-the-art WMC techniques based on pseudo-Boolean function manipulation,
  competitive with the more traditional WMC algorithms based on knowledge
  compilation and backtracking search. We present an algorithm that transforms
  WMC instances into a format based on pseudo-Boolean functions while
  eliminating around \SI{57}{\percent} of variables on average across various
  Bayesian network encodings. Moreover, we identify sufficient conditions for
  such a variable removal to be possible. Our experiments show significant
  improvement in WMC-based Bayesian network inference, outperforming the current
  state of the art.

@@

% NOTE: What's wrong with previous proposed solutions? How does mine differ?
One approach to this problem considers weighted clauses and probabilistic
semantics based on Markov networks \cite{DBLP:conf/uai/GogateD10}. However, with
a new representation comes the need to invent new encodings and inference
algorithms. Our work is similar in that it introduces a new representation for
computational problems but can reuse recent WMC algorithms based on
pseudo-Boolean function manipulation, namely, \textsf{ADDMC}
\cite{DBLP:conf/aaai/DudekPV20} and \textsf{DPMC} \cite{DBLP:conf/cp/DudekPV20}.
Furthermore, we identify sufficient conditions for transforming a WMC instance
into our new format. As many WMC inference algorithms such as
\textsf{Ace}\footnote{\url{http://reasoning.cs.ucla.edu/ace/}}, \textsf{c2d}
\cite{DBLP:conf/ecai/Darwiche04}, and \textsf{miniC2D}
\cite{DBLP:conf/ijcai/OztokD15} work by compilation to tractable representations
such as arithmetic circuits, deterministic, decomposable negation normal form
\cite{DBLP:journals/jancl/Darwiche01}, and sentential decision diagrams (SDDs)
\cite{DBLP:conf/ijcai/Darwiche11}, another way to avoid parameter variables
could be via direct compilation to a more convenient representation. However,
while direct compilation of Bayesian networks to SDDs has been investigated
\cite{DBLP:conf/ecsqaru/ChoiKD13}, SDDs are not expressive enough to avoid the
issue. Furthermore, to the best of the authors' knowledge, neither
compilation/inference technique has a publicly available implementation.

% NOTE: What are the key components of my approach and results? Also include any
% specific limitations. What about your project is new or different from
% existing solutions? List of refutable contributions (with forward references
% to the rest of the paper) What is the outcome of the project?
In this work, we introduce a way to transform WMC problems into a new format
based on pseudo-Boolean functions---\emph{pseudo-Boolean projection} (PBP). We
formally show that every WMC problem instance has a corresponding PBP instance
and identify conditions under which this transformation can remove parameter
variables. Four out of the five known WMC encodings for Bayesian networks
\cite{DBLP:conf/ecai/BartKLM16,DBLP:conf/ijcai/ChaviraD05,DBLP:conf/sat/ChaviraD06,DBLP:conf/kr/Darwiche02,DBLP:conf/aaai/SangBK05}
can indeed be simplified in this manner. This transformation enables two
encodings that were previously incompatible with most WMC algorithms (due to
using a different definition of WMC
\cite{DBLP:conf/ijcai/ChaviraD05,DBLP:conf/sat/ChaviraD06}) to be run with
\textsf{ADDMC} and \textsf{DPMC} and results in a significant performance boost
for one other encoding, allowing it to surpass the current state of the art.
Finally, our theoretical contributions result in a convenient algebraic way of
reasoning about two-valued pseudo-Boolean functions and position WMC encodings
on common ground, identifying their key properties and assumptions.

@@

\begin{definition}[WMC]
  A \emph{WMC instance} is a tuple $(\phi, X_I, X_P, w)$, where $X_I$ is
  the set of \emph{indicator} variables, $X_P$ is the set of \emph{parameter}
  variables (with $X_I \cap X_P = \emptyset$), $\phi$ is a propositional formula
  in CNF over $X_I \cup X_P$, and $w\colon X_I \cup X_P \cup \{\neg x \mid x \in
  X_I \cup X_P\} \to \mathbb{R}$ is the weight function. Its \emph{answer} is
  $\sum_{Y \models \phi} \prod_{Y \models l} w(l)$.
\end{definition}

That is, the answer to a WMC instance is the sum of the weights of all models of
$\phi$, where the weight of a model is defined as the product of the weights of
all (positive and negative) literals in it. Our definition of WMC is largely
based on the standard definition \cite{DBLP:journals/ai/ChaviraD08}, but
explicitly partitions variables into indicator and parameter variables. In
practice, we identify this partition in one of two ways. If an encoding is
generated by \textsf{Ace}, then variable types are explicitly identified in the
LMAP file generated alongside the encoding. Otherwise, we define a variable $x$
to be a parameter variable if $w(x) = w(\neg x) = 1$. Next, we formally define a
variation of the WMC problem used by some of the Bayesian network encodings
\cite{DBLP:conf/ijcai/ChaviraD05,DBLP:conf/sat/ChaviraD06}.

@@

\subsection{Bayesian Network Encodings}

@@

WMC is a well-established technique for Bayesian network inference, particularly
effective on networks where most variables have only a few possible values
\cite{DBLP:conf/kr/Darwiche02}. Many ways of encoding a Bayesian network into a
WMC instance have been proposed. We will refer to them based on the initials of
the authors and the year of publication. Darwiche was the first to suggest the
\texttt{d02} \cite{DBLP:conf/kr/Darwiche02} encoding that, in many ways, remains
the foundation behind most other encodings. He also introduced the distinction
between \emph{indicator} and \emph{parameter variables}; the former represent
variable-value pairs in the Bayesian network, while the latter are associated
with probabilities in the CPTs. The encoding \texttt{sbk05}
\cite{DBLP:conf/aaai/SangBK05} is the only encoding that deviates from this
arrangement: for each variable in the Bayesian network, one indicator variable
acts simultaneously as a parameter variable. Chavira and Darwiche propose
\texttt{cd05} \cite{DBLP:conf/ijcai/ChaviraD05} where they shift from WMC to
minimum-cardinality WMC because that allows the encoding to have fewer variables
and clauses. In particular, they propose a way to use the same parameter
variable to represent all probabilities in a CPT that are equal and keep only
clauses that `imply' parameter variables (i.e., omit clauses where a parameter
variable implies indicator variables). In their next encoding, \texttt{cd06}
\cite{DBLP:conf/sat/ChaviraD06}, the same authors optimise the aforementioned
implication clauses, choosing the smallest sufficient selection of indicator
variables. A decade later, Bart et al. present \texttt{bklm16}
\cite{DBLP:conf/ecai/BartKLM16} that improves upon \texttt{cd06} in two ways.
First, they optimise the number of indicator variables used per Bayesian network
variable from a linear to a logarithmic amount. Second, they introduce a scaling
factor that can `absorb' one probability per Bayesian network variable. However,
for this work, we choose to disable the latter improvement since this scaling
factor is often small enough to be indistinguishable from zero without the use
of arbitrary precision arithmetic. Indeed, even a small Bayesian network with
seven mutually independent binary variables, 0.1 and 0.9 probabilities each, is
already big enough for the scaling factor to be exactly equal to zero (as
produced by the \texttt{bklm16}
encoder\footnote{\url{http://www.cril.univ-artois.fr/kc/bn2cnf.html}}).

@@

\begin{definition}[PBP Instance] \label{def:new_wmc}
  A PBP instance is a tuple $(F, X, \omega)$, where $X$ is the set of variables,
  $F$ is a set of two-valued pseudo-Boolean functions $2^X \to \mathbb{R}$, and
  $\omega \in \mathbb{R}$ is the scaling factor. Its \emph{answer} is $\omega
  \cdot \left(\exists_X\prod_{f \in F}f\right)(\emptyset)$.
\end{definition}

Adding scaling factor $\omega$ to the definition allows us to remove clauses
that consist entirely of a single parameter variable. The idea of extracting
some of the structure of the WMC instance into an external multiplicative factor
was inspired by the \texttt{bklm16} encoding.

@@

In this section, we describe an algorithm for transforming WMC instances to the
PBP format while removing all parameter variables. The algorithm works on four
out of the five Bayesian network encodings. There is no obvious way to adjust it
to work with \texttt{sbk05} because the roles of indicator and parameter (i.e.,
`chance') variables overlap \cite{DBLP:conf/aaai/SangBK05}. The algorithm is
based on several observations that will be made more precise in
\cref{sec:proof}. First, all weights except for $\{w(p) \mid p \in X_P\}$ are
redundant as they either duplicate an already-defined weight or are equal to
one. Second, each clause has at most one parameter variable. Third, if the
parameter variable is negated, we can ignore the clause (this idea first appears
in the \texttt{cd05} paper \cite{DBLP:conf/ijcai/ChaviraD05}). Note that while
we formulate our algorithm as a sequel to the WMC encoding procedure primarily
because the implementations of Bayesian network WMC encodings are all
closed-source, as all transformations in the algorithm are local, it can be
efficiently incorporated into a WMC encoding algorithm with no slowdown.

@@

\begin{proposition} \label{prop:equivalence}
  Let $(\phi, X_I, X_P, w)$ be a WMC instance. Then
  \begin{equation}
  \left(\left\{[c]_0^1 \;\middle|\; c \in \phi\right\} \cup \left\{[x]_{w(\neg
        x)}^{w(x)} \;\middle|\; x \in X_I \cup X_P\right\}, X_I \cup X_P,
    1\right) \label{eq:new_wmc}
  \end{equation}
  is a PBP instance with the same answer.
\end{proposition}
\begin{proof}
  Let $f = \prod_{c \in \phi} [c]_0^1$, and $g = \prod_{x \in X_I \cup X_P}
  [x]_{w(\neg x)}^{w(x)}$. Then the WMC answer is
  \eqref{eq:new_wmc} is
  \[
    (\exists_{X_I \cup X_P} fg)(\emptyset) = \sum_{Y \subseteq X_I \cup X_P}
    (fg)(Y) = \sum_{Y \subseteq X_I \cup X_P} f(Y)g(Y)
  \]
  by \cref{lemma:sum}. Note that
  \[
    f(Y) =
    \begin{cases}
      1 & \text{if } Y \models \phi, \\
      0 & \text{otherwise},
    \end{cases}
    \quad
    \text{and}
    \quad
    g(Y) = \prod_{Y \models l} w(l),
  \]
  which means that $\sum_{Y \subseteq X_I \cup X_P} f(Y)g(Y) = \sum_{Y \models
    \phi} \prod_{Y \models l} w(l)$ as required.
\end{proof}

\begin{theorem}[Correctness for WMC] \label{thm:correctness}
  \Cref{alg:transformation}, when given a WMC instance $(\phi, X_I, X_P,
  w)$, returns PBP instance with the same answer, provided the following
  conditions are satisfied:
  \begin{enumerate}
  \item for all indicator variables $i \in X_I$, $w(i) = w(\neg i) =
    1$, \label{cond:one}
  \item and either
    \begin{enumerate}
    \item for all parameter variables $p \in X_P$, there is a non-empty family
      of literals $(l_i)_{i=1}^n$ such that \label{cond:d02}
      \begin{enumerate}
      \item $w(\neg p) = 1$,
      \item $l_i \in X_I$ or $\neg l_i \in X_I$ for all $i = 1, \dots, n$,
        \label{condition:equivalence1}
      \item and $\{ c \in \phi \mid p \in c \text{ or } \neg p \in c \} =
        \left\{p \lor \bigvee_{i=1}^n \neg l_i \right\} \cup \{ l_i \lor \neg p
        \mid i = 1, \dots, n \}$; \label{condition:equivalence2}
      \end{enumerate}
    \item or for all parameter variables $p \in X_P$, \label{cond:bklm16}
      \begin{enumerate}
      \item $w(p) + w(\neg p) = 1$,
      \item for any clause $c \in \phi$, $|c \cap X_P| \le 1$, \label{cond:2b2}
      \item there is no clause $c \in \phi$ such that $\neg p \in
        c$, \label{cond:2b3}
      \item if $\{ p \} \in \phi$, then there is no clause $c \in \phi$ such
        that $c \ne \{ p \}$ and $p \in c$, \label{cond:just_parameter}
        % 'prime implicants are logically disjoint'
      \item and for any $c, d \in \phi$ such that $c \ne d$, $p \in c$ and $p
        \in d$, $\bigwedge_{l \in c \setminus \{ p \}} \neg l \land \bigwedge_{l
          \in d \setminus \{ p \}} \neg l$ is false. \label{cond:disjoint}
      \end{enumerate}
    \end{enumerate}
  \end{enumerate}
\end{theorem}

\Cref{cond:one} applies to all four WMC encodings under consideration and
ensures that only parameter variables carry nontrivial weights. \Cref{cond:d02}
(for \texttt{d02}) simply states that each parameter variable is equivalent to a
conjunction of indicator literals. \Cref{cond:bklm16} is for encodings that have
implications rather than equivalences associated with parameter variables
(which, in this case, is \texttt{bklm16}). It ensures that each clause has at
most one positive parameter literal and no negative ones, and that at most one
implication clause per any parameter variable $p \in X_P$ can `force $p$ to be
positive'.

@@

  Let us first consider how the unfinished WMC instance $(F, X_I, \omega)$ after
  the loop on \crefrange{line:foreach1start}{line:foreach1end} differs from
  \eqref{eq:new_wmc2}. Note that \cref{alg:transformation} leaves each $c \in
  \phi_*$ unchanged, i.e., adds $[c]_0^1$ to $F$. We can then fix an arbitrary
  $p \in X_P$ and let $F_p$ be the set of functions added to $F$ as a
  replacement of $\phi_p$. It is sufficient to show that
  \begin{equation} \label{eq:to_show}
    \omega \prod_{f \in F_p} f = \exists_p [p]_{w(\neg p)}^{w(p)} \prod_{c \in \phi_p} [c]_0^1.
  \end{equation}
  Note that under \cref{cond:d02}, $\bigwedge_{c \in \phi_p} c \equiv p
  \Leftrightarrow \bigwedge_{i=1}^n l_i$ for some family of indicator variable
  literals $(l_i)_{i=1}^n$. Thus,
  \[
    \exists_p [p]_{w(\neg p)}^{w(p)} \prod_{c \in \phi_p} [c]_0^1 = \exists_p
    [p]_1^{w(p)} \left[ p \Leftrightarrow \bigwedge_{i=1}^n l_i \right]_0^1.
  \]
  If $w(p) = 1$, then
  \begin{equation} \label{eq:bigsums}
    \exists_p [p]_1^{w(p)} \left[ p \Leftrightarrow \bigwedge_{i=1}^n l_i \right]_0^1 = \left.\left[ p \Leftrightarrow \bigwedge_{i=1}^n l_i \right]_0^1\right|_{p=1} + \left.\left[ p \Leftrightarrow \bigwedge_{i=1}^n l_i \right]_0^1\right|_{p=0}.
  \end{equation}
  Since for any input, $\bigwedge_{i=1}^n l_i$ is either true or false, exactly
  one of the two summands in \cref{eq:bigsums} will be equal to one, and the
  other will be equal to zero, and so
  \[
    \left.\left[ p \Leftrightarrow \bigwedge_{i=1}^n l_i
      \right]_0^1\right|_{p=1} + \left.\left[ p \Leftrightarrow \bigwedge_{i=1}^n
        l_i \right]_0^1\right|_{p=0} = 1,
  \]
  where $1$ is a pseudo-Boolean function that always returns one. On the other
  side of \cref{eq:to_show}, since $F_p = \emptyset$, and $\omega$ is unchanged,
  we get $\omega\prod_{f \in F_p} f = 1$, and so \cref{eq:to_show} is satisfied
  under \cref{cond:d02} when $w(p) = 1$.

@@

\begin{figure}
  \centering
  \input{cumulative.tex}%
  \caption{Cumulative numbers of instances solved by each algorithm-encoding
    pair over time. The dotted line denotes the total number of instances used.}
  \label{fig:cumulative}
\end{figure}

@@

>>>>

% The abstract should briefly summarize the contents of the paper in
% 150--250 words.
% 1. Introduction. In one sentence, what’s the topic?
% 2. State the problem you tackle.
% 3. Summarize (in one sentence) why nobody else has adequately answered the research question yet.
% 4. Explain, in one sentence, how you tackled the research question.
% 5. In one sentence, how did you go about doing the research that follows from your big idea.
% 6. As a single sentence, what’s the key impact of your research?
\begin{abstract}
  Weighted model counting (WMC) is a powerful computational technique for a
  variety of problems, especially commonly used for probabilistic inference.
  However, the standard definition of WMC that puts weights on literals often
  necessitates WMC encodings to include additional variables and clauses just so
  each weight can be attached to a literal. This paper complements previous work
  by considering WMC instances in their full generality and using recent
  state-of-the-art WMC techniques based on pseudo-Boolean function manipulation,
  competitive with the more traditional WMC algorithms based on knowledge
  compilation and backtracking search. We present an algorithm that transforms
  WMC instances into a format based on pseudo-Boolean functions while
  eliminating around \SI{43}{\percent} of variables on average across various
  Bayesian network encodings. Moreover, we identify sufficient conditions for
  such a variable removal to be possible. Our experiments show significant
  improvement in WMC-based Bayesian network inference, outperforming the current
  state of the art.

@@

% NOTE: What's wrong with previous proposed solutions? How does mine differ?
One approach to this problem considers weighted clauses and probabilistic
semantics based on Markov networks \cite{DBLP:conf/uai/GogateD10}. However, with
a new representation comes the need to invent new encodings and inference
algorithms. Our work is similar in spirit in that it introduces a new
representation for computational problems but can reuse recent WMC algorithms
based on pseudo-Boolean function manipulation, namely, \textsf{ADDMC}
\cite{DBLP:conf/aaai/DudekPV20} and \textsf{DPMC} \cite{DBLP:conf/cp/DudekPV20}.
Furthermore, we identify sufficient conditions for transforming a WMC instance
into our new format. As many WMC inference algorithms such as
\textsf{Ace}, \textsf{c2d}
\cite{DBLP:conf/ecai/Darwiche04}, and \textsf{miniC2D}
\cite{DBLP:conf/ijcai/OztokD15} work by compilation to tractable representations
such as arithmetic circuits, deterministic, decomposable negation normal form
\cite{DBLP:journals/jancl/Darwiche01}, and sentential decision diagrams (SDDs)
\cite{DBLP:conf/ijcai/Darwiche11}, another way to avoid parameter variables
could be via direct compilation to a more convenient representation. Direct
compilation of Bayesian networks to SDDs has been investigated
\cite{DBLP:conf/ecsqaru/ChoiKD13}. However, SDDs only support weights on
literals, and so are not expressive enough to avoid the issue. To the best of
the authors' knowledge, neither approach
\cite{DBLP:conf/ecsqaru/ChoiKD13,DBLP:conf/uai/GogateD10} has a publicly
available implementation.

% NOTE: What are the key components of my approach and results? Also include any
% specific limitations. What about your project is new or different from
% existing solutions? List of refutable contributions (with forward references
% to the rest of the paper) What is the outcome of the project?
In this work, we introduce a way to transform WMC problems into a new format
based on pseudo-Boolean functions---\emph{pseudo-Boolean projection} (PBP). We
formally show that every WMC problem instance has a corresponding PBP instance
and identify conditions under which this transformation can remove parameter
variables. Four out of the five known WMC encodings for Bayesian networks
\cite{DBLP:conf/ecai/BartKLM16,DBLP:conf/ijcai/ChaviraD05,DBLP:conf/sat/ChaviraD06,DBLP:conf/kr/Darwiche02,DBLP:conf/aaai/SangBK05}
can indeed be simplified in this manner. We are able to eliminate
\SI{43}{\percent} of variables on average and up to \SI{99}{\percent} on some
instances. This transformation enables two encodings that were previously
incompatible with most WMC algorithms (due to using a different definition of
WMC \cite{DBLP:conf/ijcai/ChaviraD05,DBLP:conf/sat/ChaviraD06}) to be run with
\textsf{ADDMC} and \textsf{DPMC} and results in a significant performance boost
for one other encoding, making it about three times faster than the state of the
art. Finally, our theoretical contributions result in a convenient algebraic way
of reasoning about two-valued pseudo-Boolean functions and position WMC
encodings on common ground, identifying their key properties and assumptions.

@@

\begin{definition}[WMC]
  A \emph{WMC instance} is a tuple $(\phi, X_I, X_P, w)$, where $X_I$ is
  the set of \emph{indicator variables}, $X_P$ is the set of \emph{parameter
  variables} (with $X_I \cap X_P = \emptyset$), $\phi$ is a propositional formula
  in CNF over $X_I \cup X_P$, and $w\colon X_I \cup X_P \cup \{\neg x \mid x \in
  X_I \cup X_P\} \to \mathbb{R}$ is a \emph{weight function} such that $w(x) =
  w(\neg x) = 1$ for all $x \in X_I$. The \emph{answer} of the instance is
  $\sum_{Y \models \phi} \prod_{Y \models l} w(l)$.
\end{definition}

That is, the answer to a WMC instance is the sum of the weights of all models of
$\phi$, where the weight of a model is defined as the product of the weights of
all (positive and negative) literals in it. Our definition of WMC is largely
based on the standard definition \cite{DBLP:journals/ai/ChaviraD08}, but
explicitly partitions variables into indicator and parameter variables. In
practice, we identify this partition in one of two ways. If an encoding is
generated by \textsf{Ace}\footnote{\textsf{Ace}
  \cite{DBLP:journals/ai/ChaviraD08} implements most of the Bayesian network
  encodings and can also be used for compilation (and thus inference). It is
  available at \url{http://reasoning.cs.ucla.edu/ace/}.}, then variable types
are explicitly identified in a file generated alongside the encoding. Otherwise,
we take $X_I$ to be the set of all variables $x$ such that $w(x) = w(\neg x) =
1$. Next, we formally define a variation of the WMC problem used by some of the
Bayesian network encodings
\cite{DBLP:conf/ijcai/ChaviraD05,DBLP:conf/sat/ChaviraD06}.

@@

\begin{example}
  Let $\phi = (x \lor y) \land (\neg x \lor \neg y) \land (\neg x \lor p) \land
  (\neg y \lor q) \land x$, $X_I = \{ x, y \}$, $X_P = \{ p, q \}$, $w(p) =
  0.2$, $w(q) = 0.8$, and $w(\neg p) = w(\neg q) = 1$. Then $\phi$ has two
  models: $\{x, p\}$ and $\{ x, p, q \}$ with weights $0.2$ and $0.2 \times 0.8
  = 0.16$, respectively. The WMC answer is then $0.2 + 0.16 = 0.36$, and the
  minimum-cardinality WMC answer is $0.2$.
\end{example}

\subsection{Bayesian Network Encodings} \label{sec:encodings}

@@

WMC is a well-established technique for Bayesian network inference, particularly
effective on networks where most variables have only a few possible values
\cite{DBLP:conf/kr/Darwiche02}. Many ways of encoding a Bayesian network into a
WMC instance have been proposed. We will refer to them based on the initials of
the authors and the year of publication. Darwiche was the first to suggest the
\texttt{d02} \cite{DBLP:conf/kr/Darwiche02} encoding that, in many ways, remains
the foundation behind most other encodings. He also introduced the distinction
between \emph{indicator} and \emph{parameter variables}; the former represent
variable-value pairs in the Bayesian network, while the latter are associated
with probabilities in the CPTs. The encoding \texttt{sbk05}
\cite{DBLP:conf/aaai/SangBK05} is the only encoding that deviates from this
arrangement: for each variable in the Bayesian network, one indicator variable
acts simultaneously as a parameter variable. Chavira and Darwiche propose
\texttt{cd05} \cite{DBLP:conf/ijcai/ChaviraD05} where they shift from WMC to
minimum-cardinality WMC because that allows the encoding to have fewer variables
and clauses. In particular, they propose a way to use the same parameter
variable to represent all probabilities in a CPT that are equal and keep only
clauses that `imply' parameter variables (i.e., omit clauses where a parameter
variable implies indicator variables). In their next encoding, \texttt{cd06}
\cite{DBLP:conf/sat/ChaviraD06}, the same authors optimise the aforementioned
implication clauses, choosing the smallest sufficient selection of indicator
variables. A decade later, Bart et al. present \texttt{bklm16}
\cite{DBLP:conf/ecai/BartKLM16} that improves upon \texttt{cd06} in two ways.
First, they optimise the number of indicator variables used per Bayesian network
variable from a linear to a logarithmic amount. Second, they introduce a scaling
factor that can `absorb' one probability per Bayesian network variable. However,
for this work, we choose to disable the latter improvement since this scaling
factor is often small enough to be indistinguishable from zero without the use
of arbitrary precision arithmetic. Indeed, even a small Bayesian network with
seven mutually independent binary variables, 0.1 and 0.9 probabilities each, is
already big enough for the scaling factor to be exactly equal to zero (as
produced by the \texttt{bklm16}
encoder\footnote{\url{http://www.cril.univ-artois.fr/kc/bn2cnf.html}}).
% TODO (on encodings and clauses): I realise the paper is already long but it
% might be useful to have a quick example
% TODO: If I am reading this correctly you are making a stark statement about
% the inefficacy. It might be good to rephrase this differently. You might want
% to say that there are many corner cases where there are no improvements. 
% TODO: Is there a reference for this example? If yes provide one. If no, maybe
% you could provide a link to the output, or alternatively, you could say
% something like the reader is free to check that such an example produces no
% improvements
% TODO: is this example the same as the previous one?

@@

\begin{definition}[PBP Instance] \label{def:new_wmc}
  A PBP instance is a tuple $(F, X, \omega)$, where $X$ is the set of variables,
  $F$ is a set of two-valued pseudo-Boolean functions $2^X \to \mathbb{R}$, and
  $\omega \in \mathbb{R}$ is the scaling factor.\footnote{
    Adding scaling factor $\omega$ to the definition allows us to remove clauses
    that consist entirely of a single parameter variable. The idea of extracting
    some of the structure of the WMC instance into an external multiplicative factor
    was loosely inspired by the \texttt{bklm16} encoding, where it is used
    to subsume the most commonly occurring probability of each CPT
    \cite{DBLP:conf/ecai/BartKLM16}.} Its \emph{answer} is $\omega \cdot
  \left(\exists_X\prod_{f \in F}f\right)(\emptyset)$.
\end{definition}

@@

In this section, we describe an algorithm for transforming WMC instances to the
PBP format while removing all parameter variables. The algorithm works on four
out of the five Bayesian network encodings: \texttt{bklm16}
\cite{DBLP:conf/ecai/BartKLM16}, \texttt{cd05}
\cite{DBLP:conf/ijcai/ChaviraD05}, \texttt{cd06}
\cite{DBLP:conf/sat/ChaviraD06}, and \texttt{d02}
\cite{DBLP:conf/kr/Darwiche02}. There is no obvious way to adjust it to work
with \texttt{sbk05} because the roles of indicator and parameter (i.e.,
`chance') variables overlap \cite{DBLP:conf/aaai/SangBK05}. The algorithm is
based on several observations that will be made more precise in
\cref{sec:proof}. First, all weights except for $\{w(p) \mid p \in X_P\}$ are
redundant as they either duplicate an already-defined weight or are equal to
one. Second, each clause has at most one parameter variable. Third, if the
parameter variable is negated, we can ignore the clause (this idea first appears
in the \texttt{cd05} paper \cite{DBLP:conf/ijcai/ChaviraD05}). Note that while
we formulate our algorithm as a sequel to the WMC encoding procedure primarily
because the implementations of Bayesian network WMC encodings are all
closed-source, as all transformations in the algorithm are local, it can be
efficiently incorporated into a WMC encoding algorithm with no slowdown.

@@

\begin{proposition} \label{prop:equivalence}
  Let $(\phi, X_I, X_P, w)$ be a WMC instance. Then
  \begin{equation}
  \left(\left\{[c]_0^1 \;\middle|\; c \in \phi\right\} \cup \left\{[x]_{w(\neg
        x)}^{w(x)} \;\middle|\; x \in X_I \cup X_P\right\}, X_I \cup X_P,
    1\right) \label{eq:new_wmc}
  \end{equation}
  is a PBP instance with the same answer.
\end{proposition}
% TODO: It’s a bit strange to say same answer, when you don’t precisely refer to
% this anywhere. One option is to say the above is the WMC instance and let
% Delta be the weighted model count. Then the new encoding Has a weighted model
% count of delta as well. Or something like that.
% But looking at the text below, I see that you use this phrasing everywhere,
% and I admit my formulation is awkward. So maybe the best compromise here is
% just simply add a footnote explaining what exactly you mean when you say same
% answer
% TODO: If it is the case that you cannot include the proofs in line, and indeed
% it might be better for readability to have an appendix with all the proofs, it
% might make sense to provide the type of argument you are using in the proofs
% so that readers know what to expect when they look into it
\begin{proof}
  Let $f = \prod_{c \in \phi} [c]_0^1$, and $g = \prod_{x \in X_I \cup X_P}
  [x]_{w(\neg x)}^{w(x)}$. Then the WMC answer is
  \eqref{eq:new_wmc} is
  \[
    (\exists_{X_I \cup X_P} fg)(\emptyset) = \sum_{Y \subseteq X_I \cup X_P}
    (fg)(Y) = \sum_{Y \subseteq X_I \cup X_P} f(Y)g(Y)
  \]
  by \cref{lemma:sum}. Note that
  \[
    f(Y) =
    \begin{cases}
      1 & \text{if } Y \models \phi, \\
      0 & \text{otherwise},
    \end{cases}
    \quad
    \text{and}
    \quad
    g(Y) = \prod_{Y \models l} w(l),
  \]
  which means that $\sum_{Y \subseteq X_I \cup X_P} f(Y)g(Y) = \sum_{Y \models
    \phi} \prod_{Y \models l} w(l)$ as required.
\end{proof}

\begin{theorem}[Correctness for WMC] \label{thm:correctness}
  \Cref{alg:transformation}, when given a WMC instance $(\phi, X_I, X_P,
  w)$, returns a PBP instance with the same answer, provided \emph{either} of
  the two conditions is satisfied:
  \begin{enumerate}
  \item for all $p \in X_P$, there is a non-empty family of literals
    $(l_i)_{i=1}^n$ such that \label{cond:d02}
    \begin{enumerate}
    \item $w(\neg p) = 1$,
    \item $l_i \in X_I$ or $\neg l_i \in X_I$ for all $i = 1, \dots, n$,
      \label{condition:equivalence1}
    \item and $\{ c \in \phi \mid p \in c \text{ or } \neg p \in c \} =
      \left\{p \lor \bigvee_{i=1}^n \neg l_i \right\} \cup \{ l_i \lor \neg p
      \mid i = 1, \dots, n \}$; \label{condition:equivalence2}
    \end{enumerate}
  \item or for all $p \in X_P$, \label{cond:bklm16}
    \begin{enumerate}
    \item $w(p) + w(\neg p) = 1$,
    \item for any clause $c \in \phi$, $|c \cap X_P| \le 1$, \label{cond:2b2}
    \item there is no clause $c \in \phi$ such that $\neg p \in
      c$, \label{cond:2b3}
    \item if $\{ p \} \in \phi$, then there is no clause $c \in \phi$ such
      that $c \ne \{ p \}$ and $p \in c$, \label{cond:just_parameter}
    \item and for any $c, d \in \phi$ such that $c \ne d$, $p \in c$ and $p
      \in d$, $\bigwedge_{l \in c \setminus \{ p \}} \neg l \land \bigwedge_{l
        \in d \setminus \{ p \}} \neg l$ is false. \label{cond:disjoint}
    \end{enumerate}
  \end{enumerate}
\end{theorem}

\Cref{cond:d02} (for \texttt{d02}) simply states that each parameter variable is
equivalent to a conjunction of indicator literals. \Cref{cond:bklm16} is for
encodings that have implications rather than equivalences associated with
parameter variables (which, in this case, is \texttt{bklm16}). It ensures that
each clause has at most one positive parameter literal and no negative ones, and
that at most one implication clause per any parameter variable $p \in X_P$ can
`force $p$ to be positive'.

@@

  Let us first consider how the unfinished WMC instance $(F, X_I, \omega)$ after
  the loop on \crefrange{line:foreach1start}{line:foreach1end} differs from
  \eqref{eq:new_wmc2}. Note that \cref{alg:transformation} leaves each $c \in
  \phi_*$ unchanged, i.e., adds $[c]_0^1$ to $F$. We can then fix an arbitrary
  $p \in X_P$ and let $F_p$ be the set of functions added to $F$ as a
  replacement of $\phi_p$. It is sufficient to show that
  \begin{equation} \label{eq:to_show}
    \omega \prod_{f \in F_p} f = \exists_p [p]_{w(\neg p)}^{w(p)} \prod_{c \in \phi_p} [c]_0^1.
  \end{equation}
  Note that under \cref{cond:d02}, $\bigwedge_{c \in \phi_p} c \equiv p
  \Leftrightarrow \bigwedge_{i=1}^n l_i$ for some family of indicator variable
  literals $(l_i)_{i=1}^n$. Thus, $\exists_p [p]_{w(\neg p)}^{w(p)} \prod_{c \in
    \phi_p} [c]_0^1 = \exists_p [p]_1^{w(p)} \left[ p \Leftrightarrow
    \bigwedge_{i=1}^n l_i \right]_0^1$. If $w(p) = 1$, then
  \begin{equation} \label{eq:bigsums}
    \exists_p [p]_1^{w(p)} \left[ p \Leftrightarrow \bigwedge_{i=1}^n l_i \right]_0^1 = \left.\left[ p \Leftrightarrow \bigwedge_{i=1}^n l_i \right]_0^1\right|_{p=1} + \left.\left[ p \Leftrightarrow \bigwedge_{i=1}^n l_i \right]_0^1\right|_{p=0}.
  \end{equation}
  Since for any input, $\bigwedge_{i=1}^n l_i$ is either true or false, exactly
  one of the two summands in \cref{eq:bigsums} will be equal to one, and the
  other will be equal to zero, and so
  \[
    \left.\left[ p \Leftrightarrow \bigwedge_{i=1}^n l_i
      \right]_0^1\right|_{p=1} + \left.\left[ p \Leftrightarrow \bigwedge_{i=1}^n
        l_i \right]_0^1\right|_{p=0} = 1,
  \]
  where $1$ is a pseudo-Boolean function that always returns one. On the other
  side of \cref{eq:to_show}, since $F_p = \emptyset$, and $\omega$ is unchanged,
  we get $\omega\prod_{f \in F_p} f = 1$, and so \cref{eq:to_show} is satisfied
  under \cref{cond:d02} when $w(p) = 1$.

@@

\begin{figure}
  \centering
  \input{cumulative.tex}%
  \caption{Cactus plot of all algorithm-encoding pairs. The dotted line
    denotes the total number of instances used.}
  \label{fig:cumulative}
\end{figure}

@@

% TODO: consider moving (some) proofs to the appendix or at least making a clear
% argument why they should be in the text.
